
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#Summary proportion for overall proportion and severe infections

# Calling necessary libraries
```{r, results='hide', message=FALSE, warning=FALSE}
library("readxl")
library(dplyr)
library(dmetar)
theme_set(theme_bw())
library(metafor)
library(meta)
library(dplyr)
library(tidyr)
library(caret)
library(hrbrthemes)
library(viridis)
library(gridExtra)
library(ggplot2)
library(grid)
library(reshape2)
library(viridis)
require(ggpubr)
require(tidyverse)
require(Hmisc)
require(corrplot)
library(lemon)
library(kableExtra)
library(data.table)
```


## Reading the dataset

```{r Reading in data}
Infection_df <- read_excel("database_trials_totaal_14nov_2021.xlsx")
```

```{r Changing column names}
setnames(Infection_df, "Patients_n...21", "n")
setnames(Infection_df, "Infections_total_n_patients", "total_infection")
setnames(Infection_df, "SIE_n_patients", "severe_infection")
```

```{r function to compute the mode}
getmode <- function(vec) {
  #function to compute the mode for a 1-D array or vector
   uniqv <- unique(vec)
   uniqv[which.max(tabulate(match(vec, uniqv)))]
}
```


```{r aggregating selected columns by PMID}
# We aggregated n, pneumonia infection, authors, year, follow-up week, study design, and severe infection by PMID
data.agg <- Infection_df %>% 
  group_by(PMID) %>% 
  summarise(Infection = sum(total_infection, na.rm = TRUE),
          n = sum(n, na.rm = TRUE) ,
          Pneumonia_infection = sum(Pneumonia_patients, na.rm = T),
          Authors = getmode(Author),
          Year = getmode(Year),
          Follow_up_weeks = getmode(Follow_up_duration_weeks),
          studydesign = getmode(Design),            
          SIE_infection = sum(SIE_min_estimate_patients, na.rm = TRUE))      
```


#Grouping follow-up length into three <=12 weeks, 13 - 26 weeks, and >26 weeks 
```{r recoding follow-up weeks}
#Grouping follow-up week
followup.length.group <- NULL
for ( i in seq(1,dim(Infection_df)[1], by=1)){
  if(Infection_df$Follow_up_duration_weeks[i] <= 12){followup.length.group <- append(followup.length.group, "<=12 weeks" )}
  
  if(Infection_df$Follow_up_duration_weeks[i] > 12 & Infection_df$Follow_up_duration_weeks[i] <= 26) {followup.length.group <- append(followup.length.group,  "13 - 26 weeks")}
  
  if(Infection_df$Follow_up_duration_weeks[i] > 26 & Infection_df$Follow_up_duration_weeks[i] != 999){followup.length.group <- append(followup.length.group,">26 weeks")}
  
  if(Infection_df$Follow_up_duration_weeks[i] == 999){followup.length.group <- append(followup.length.group,"NaN")}
}
```

```{r adding follow-up group to the dataframe}
#adding grouped follow up length to the existing data frame
Infection_df$followup.length.group <- followup.length.group
```


```{r combinig follow-up period and study design }
design_followup <- paste0(Infection_df$Design, Infection_df$followup.length.group)
Infection_df$design_followup <- design_followup
```


#Overall proportion
```{r removing rows with missing proportion - overall proportion}
#In the dataset 999 was put to indicate missing values in overall infection and severe infection 
Infection_df_overall <- copy(Infection_df)
Infection_df_overall$P <- Infection_df_overall$total_infection/Infection_df_overall$n
Infection_df_overall <- Infection_df_overall[!is.na(Infection_df_overall$P) & (Infection_df_overall$P  < 1) ,]
Infection_df_overall <- Infection_df_overall[Infection_df_overall$total_infection != 999,]

```

```{r computing the overall proportion}
#Computing the proportion for overall infection 
overall.logit <- escalc(xi=total_infection, ni=n, measure="PLO", data=Infection_df_overall)
overall.rma <- rma(yi,vi,data=overall.logit,method="REML", weighted=TRUE)
overall.prop.pre <-  predict(overall.rma, transf=transf.ilogit)
```


```{r printing the computed summary proportion, confidence interval, and heterogeneity }
#Print summary proportion and heterogeneity information
summary(overall.rma);print(overall.prop.pre)
```


#Serious infection
```{r removing rows with missing proportion - overall proportion}
#missing values in the serious infection are represented by 999 
Infection_df.severe <- copy(Infection_df)
Infection_df.severe$P_severe <- Infection_df.severe$severe_infection/Infection_df.severe$n
Infection_df.complete.severe <- Infection_df.severe[!is.na(Infection_df.severe$P_severe) & (Infection_df.severe$P_severe  < 1),]
 Infection_df.complete.severe <- Infection_df.complete.severe[as.integer(Infection_df.complete.severe$severe_infection) != 999,]
```

```{r computing severe infection proportion and heterogeneity indicators}
#Computing the proportion for overall infection
severe.logit <- escalc(xi=severe_infection, ni=n, measure="PLO", data=Infection_df.complete.severe)
severe.rma <- rma(yi,vi,data=severe.logit,method="REML", weighted=TRUE)
severe.prop.pre <- predict(severe.rma,transf=transf.ilogit)
```

```{r printing summary proportion and heterogeneity}
#printing the estimated proportion and rma(heterogeneity information)
print(severe.rma);print(severe.prop.pre)
```


#Forest plot overall infection

```{r randomly taking 50 studies for forest plot - severe infection}
#Preparing data for forest plot by taking 50 samples and sorting them
data.agg$p <- data.agg$Infection/data.agg$n
overall.agg.forest <- data.agg[between(data.agg$p,0,1),]
overall.agg.forest <- overall.agg.forest[sample(nrow(overall.agg.forest), 50), ]
overall.agg.forest <- overall.agg.forest[order(overall.agg.forest$p),]
```


```{r computing heterogeneity indicator for the forest plot}
#Computing the proportion for overall infection
overall.forest.logit <- escalc(xi=Infection, ni=n, measure="PLO", data=overall.agg.forest)
overall.forest.rma <- rma(yi,vi,data=overall.forest.logit,method="REML", weighted=TRUE)
#overall.forest.prop.pre <- predict(overall.forest.rma,transf=transf.ilogit)
```



```{r making forest plot for overall proportion}
overall.forest.summary=metaprop(as.integer(Infection),
                     n,
                     PMID,
                     data=overall.agg.forest,
                     #byvar=studydesign,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=TRUE,
                     tau.preset=sqrt(overall.forest.rma$tau2),
                     title="")
png("Overall_Forestplotmeta.png",width=1000,height=1000)
forest(overall.forest.summary,
       xlim = c(0,1), pscale =1, 
       rightcols=c("effect", "ci"),
       rightlabs=c("Proportion", "95% C.I."),
       leftcols = c("studlab", "event", "n"), 
       leftlabs = c("Study(PMID)", "Cases", "Total"),
       xlab = "Prevalence", 
       fs.xlab=12,
       fs.study=12,
       fs.study.lables=12,
       fs.heading=12,
       squaresize = 0.5, col.square="navy", col.square.lines="navy",
       col.diamond="navy", col.diamond.lines="navy",
       comb.fixed=FALSE,
       lty.fixed=0,
       lty.random=2, 
       type.study="square",
       type.random="diamond",
       ff.fixed="bold.italic",
       ff.random="bold.italic",
       hetlab = "Heterogeneity:",
       fs.hetstat=10,
       smlab="",
       print.Q=TRUE,
       print.pval.Q=TRUE,
       print.I2=TRUE,
       print.tau2=FALSE,
       col.by="grey",
       digits=4)
dev.off()
```

#Forest plot for Severe infection

```{r randomly taking 50 studies for forest plot - severe infection}
#Preparing data for forest plot by taking 50 samples and sorting them
data.agg$p_severe <- data.agg$SIE_infection/data.agg$n
severe.agg.forest <- data.agg[between(data.agg$p_severe,0,1),]
severe.agg.forest <- severe.agg.forest[sample(nrow(severe.agg.forest), 50), ]
severe.agg.forest <- severe.agg.forest[order(severe.agg.forest$p),]
```


```{r}
#Computing the proportion for severe infection
severe.forest.logit <- escalc(xi=SIE_infection, ni=n, measure="PLO", data=severe.agg.forest)
severe.forest.rma <- rma(yi,vi,data=severe.forest.logit,method="REML", weighted=TRUE)
#overall.forest.prop.pre <- predict(overall.forest.rma,transf=transf.ilogit)
```


```{r making forest plot for severe infection}
severe.forest.summary <- metaprop(as.integer(SIE_infection),
                     n,
                     PMID,
                     data=severe.agg.forest,
                     #byvar=studydesign,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=TRUE,
                     tau.preset=sqrt(severe.forest.rma$tau2),
                     title="")
png("Severe_Forestplotmeta.png",width=1000,height=1000)
forest(severe.forest.summary,
       xlim = c(0,1), pscale =1, 
       rightcols=c("effect", "ci"),
       rightlabs=c("Proportion", "95% C.I."),
       leftcols = c("studlab", "event", "n"), 
       leftlabs = c("Study(PMID)", "Cases", "Total"),
       xlab = "Prevalence", 
       fs.xlab=12,
       fs.study=12,
       fs.study.lables=12,
       fs.heading=12,
       squaresize = 0.5, col.square="navy", col.square.lines="navy",
       col.diamond="navy", col.diamond.lines="navy",
       comb.fixed=FALSE,
       lty.fixed=0,
       lty.random=2, 
       type.study="square",
       type.random="diamond",
       ff.fixed="bold.italic",
       ff.random="bold.italic",
       hetlab = "Heterogeneity:",
       fs.hetstat=10,
       smlab="",
       print.Q=TRUE,
       print.pval.Q=TRUE,
       print.I2=TRUE,
       print.tau2=FALSE,
       col.by="grey",
       digits=4)
dev.off()
```


#Comparing the infection rate by follow up groups

```{r computing overall proportion by study design group}
metaprop(as.integer(total_infection),
                     n,
                     PMID,
                     data=Infection_df_overall,
                     byvar=followup.length.group,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(overall.rma$tau2),
                     title="")
```


#Comparing the infection rate by follow up groups(for severe infection)

```{r comparing severe infection by study design group}
metaprop(as.integer(severe_infection),
                     n,
                     PMID,
                     data=Infection_df.complete.severe,
                     byvar=followup.length.group,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(severe.rma$tau2),
                     title="")
```


#overall proportion by study designs
```{r computing overall proportion by study design}
prop.design.overall <- metaprop(as.integer(total_infection),
                     n,
                     PMID,
                     data=Infection_df_overall,
                     byvar=Design,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(all.rma$tau2),
                     title="")
prop.design.overall
```


# Severe infection by study design
```{r}
prop.design.severe <- metaprop(as.integer(severe_infection),
                     n,
                     PMID,
                     data=Infection_df.complete.severe,
                     byvar=Design,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(overall.rma$tau2),
                     title="")
prop.design.severe
```



#overall proportion by design and follow-up group 
```{r computing overall proportion by design and follow-up group}
prop.design_followup.overall <- metaprop(as.integer(total_infection),
                     n,
                     PMID,
                     data=Infection_df_overall,
                     byvar=design_followup,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(all.rma$tau2),
                     title="")
prop.design_followup.overall

```


#Severe infection by design and follow-up group
```{r}
prop.design_followup.severe <- metaprop(as.integer(severe_infection),
                     n,
                     PMID,
                     data=Infection_df.complete.severe,
                     byvar=design_followup,
                     sm="PRAW",
                     method.ci="NAsm",
                     method.tau="DL",
                     incr=0.5,
                     allincr=FALSE,
                     addincr=FALSE,
                     tau.common=FALSE,
                     tau.preset=sqrt(severe.rma$tau2),
                     title="")
prop.design_followup.severe

```









